\mythesistitleNOTSYNPOSIS

### 2011-12-22 src/thesis.tex
Reorg-ed lit. Added themes - set sys, canonization, certifying algo - to toc.

\noindent 
\begin{tabbing}
xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
\textbf{COP}   \> Consecutive-Ones Property \\
\textbf{COT}   \> Consecutive-Ones Testing \\
\textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
\textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\
% ADD MORE HERE
\end{tabbing}

@@

\tnote{ The main themes of the thesis that must not be left out:\\
  connection of COP to ||set systems|| and thus to general
  data bases. cite:nsnrs09\\
  ||graph isomorphism|| by canonization cite:kklv10\\
  ||certificate for a problem|| cite:mcc04\\
  Also see purple book notes.  }

@@

\section{Consecutive-Ones Property Testing - a Brief Survey} %{Background}
\label{sec:background}
%\vspace{\secfirstparatrim}
In this section, a brief survey of the consecutive-ones problem and its
optimization problems is presented.

@@

@@

% \subsection{Overlap Graphs and Marginal Hyperedges} 
% An {\em
%   overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph such
% that its vertex set has a bijection to $\cF$ and there exists an edge
% between two of its vertices \iff their corresponding hyperedges
% overlap. Two hyperedges $S$ and $S'$ are said to {\em overlap},
% denoted by $S \overlap S'$, if they have a non-empty intersection and
% neither is contained in the other; $S \overlap S' \text{ iff } S \cap
% S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$
% is a spanning subgraph of $\bI(\cF)$ and not necessarily
% connected. Each connected component of $\bO(\cF)$ is called an {\em
%   overlap component}.
% % If there are $d$ overlap components in $\bO(\cF)$, the set
% % subsystems are denoted by $\cO_1, \cO_2, \ldots \cO_d$. Clearly
% % $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$, it can be
% % verified that one of the following is true.
% % \begin{enumerate}[a) ]
% % \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
% % \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
% % \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
% % \end{enumerate}
% % \vspace{\prelimspace} \xnoindent
% A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
% \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
% \cite{kklv10}. Additionally, if $S$ is such that it is contained in no
% other marginal hyperedge in $\cF$, then it is called {\em
%   super-marginal}.
% % i.e., it is inclusion maximal then it is called {\em
% %   super-marginal}.

@@

\chapter{Consecutive-Ones property - Important Results from Literature}
\tnote{TBD: have a few lines about organization of chapter}

@@

\chapter{Preliminaries to new results}
%\section{Preliminaries} 
\label{sec:prelims}

\tnote{TBD: have a few lines about organization of chapter}

This section states definitions and basic facts necessary in the scope
of this document.

\def\xnoindent {}

@@

\xnoindent The graph $T$ represents a {\em target tree} with same
number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
%\vspace{\prelimspace}

% \xnoindent
% A set system $\cF$ can be alternatively represented by a {\em
%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and hyperedges
% are the sets in $\cF$. This is a known representation for interval
% systems in literature \cite{bls99,kklv10}.  We extend this definition here to
% path systems.  
\xnoindent
A set system $\cF$ can be alternatively represented by a {\em
  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
are the sets in $\cF$. This is a known representation for interval
systems in literature \cite{bls99,kklv10}.  We extend this definition here to
path systems. Due to the equivalence of set system and hypergraph in
the scope of this paper, we drop the
subscript $_H$ in the notation and refer to both the structures by $\cF$.
%\vspace{\prelimspace}

\xnoindent Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
  isomorphic} to each other, denoted by $\cF' \cong \cF''$, iff there
exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such
that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in
$\cF'$ iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x
\in A\}$ \cite{kklv10}. This is called {\em hypergraph
  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
  required }
%\vspace{\prelimspace}

\xnoindent The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
there exists an edge between two vertices iff their corresponding
hyperedges have a non-empty intersection \cite{mcg04}.
%\vspace{\prelimspace}

@@

\xnoindent If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
is called a {\em path hypergraph} and $\cP$ is called {\em path
  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
\rightarrow V(T)$, then it is clear that there is an {\em induced path
labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
$supp(\cP) = V(T)$.
%\vspace{\prelimspace}

@@

\xnoindent
A path labeling $(\cF, \cl)$ is defined to be {\em
  feasible} if
$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
%\vspace{\prelimspace}

\xnoindent An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
a graph such that its vertex set has a bijection to $\cF$ and there
exists an edge between two of its vertices iff their corresponding
hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
  overlap}, denoted by $S \overlap S'$, if they have a non-empty
intersection and neither is contained in the other; $S \overlap S'
\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
necessarily connected. Each connected component of $\bO(\cF)$ is
called an {\em overlap component}.
% If there are $d$ overlap components
% in $\bO(\cF)$, the set subsystems are denoted by $\cO_1, \cO_2, \ldots
% \cO_d$. Clearly $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$,
% it can be verified that one of the following is true.
% \begin{enumerate}[a) ]
% \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
% \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
% \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
% \end{enumerate}
%\vspace{\prelimspace}

\xnoindent
A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
{\em super-marginal}.

\xnoindent A {\em star} graph is a complete bipartite graph $K_{1,p}$
which is clearly a tree and $p$ is the number of leaves. The vertex
with maximum degree is called the {\em center} of the star and the
edges are called {\em rays} of the star. A {\em $k$-subdivided star}
is a star with all its rays subdivided exactly $k$ times. The
definition of a {\em ray of a $k$-subdivided star} is extended to the path
from the center to a leaf. It is clear that all rays are of length $k+2$.

@@

\tnote{TBD: have a few lines about organization of chapter}
The necessary preliminaries with definitions
etc. are presented in
Section~\ref{sec:prelims}. Section~\ref{sec:feasible} documents the
characterization of a feasible path labeling and finally,
Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
find the tree path labeling of a given set system from a given
$k$-subdivided tree.

@@

\def \xnoindent  {}

\xnoindent In the remaining part of this section we show that $(\cF,
\cl)$ is feasible if and only if it is an ICPPL and
Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two levels of
filtering of $(\cF, \cl)$ to make it simpler while retaining the set
of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
% One direction of this claim isclear: that if a path
% labeling is feasible, then all intersection cardinalities are
% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL iff input PL
%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
% the sets map to paths, ``preserving'' the path labeling.  The rest
% of the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
its correctness.  This algorithm refines the path labeling by
processing pairs of paths in $\cF^\cl$ that share a leaf until no two
paths in the new path labeling share any leaf.

@@

  \xnoindent
 \begin{enumerate}[\xbullet]
  \item {\em Invariant} I/II
    \begin{enumerate}[{I/II}a $|$] % \textbullet 
    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
      trivially true by induction hypothesis.
    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
      $\cF_{j-1}$, then it must be one of the new sets added in
      $\cF_j$. In this case, it is clear that for each new set, the
      image under $\cl_j$ is a path since by definition the chosen
      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
      common leaf. Thus invariant I is proven.\\
      Moreover, due to induction hypothesis of invariant III and the
      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
      indeed true in the $j$th iteration for any of the new sets.  If
      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
      |\cl_j(S_1 \setminus S_2)|
      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
    \end{enumerate}
  \item {\em Invariant} III
    \begin{enumerate}[{III}a $|$]
    \item {\em $R$ and $R'$ are not new sets.} It is in
      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
      induction hypothesis, Lemma~\ref{lem:setminuscard} and
      definition of $\cl_j$, it follows that invariant III is true no
      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
      S_1$. Note $R'$ is not a new set.\\

@@

\xnoindent As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
is such that there is exactly one set in $\cF$ with $v$ as a vertex in
the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
elements in $supp(\cF)$ whose images are leaves in a hypergraph
isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
  \ne S} \hS$. In other words, $x$ is an element present in no other
set in $\cF$ except $S$. This is intuitive since $v$ is present in no
other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
leaf $v$ are then removed from the set $S$ and path $\cl(S)$
respectively. After doing this for all leaves in $T$, all path images
in the new path labeling $\cl_{new}$ except leaf labels (a path that
has only a leaf is called the {\em leaf label} for the corresponding
single element hyperedge or set) are paths from a new pruned tree $T_0
= T \setminus \{v \mid v \text{ is a leaf in }
T\}$. Algorithm~\ref{leafasgn} is now presented with details.

@@

\xnoindent
Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
loop. This will abort our procedure of finding the hypergraph
isomorphism. The following lemma shows that this will not happen.

@@

\xnoindent We have seen two filtering algorithms above, namely,
Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
{\tt filter\_2} which when executed serially repectively result in a
new ICPPL on the same universe $U$ and tree $T$. We also proved that
if the input is indeed feasible, these algorithms do indeed output the
filtered ICPPL. Now we present the algorithmic characterization of a
feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.

\xnoindent Algorithm~\ref{al:icppl-find-isomorph} computes a
hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
and Algorithm~\ref{leafasgn} and pruning the leaves of the input
tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
leaf labels are removed from the path labeling $\cl_2$ and the
corresponding elements are removed from the corresponding sets in
$\cF_2$. This tree pruning algorithm is recursively called on the
altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
recursive call returns the bijection $\phi''$ for the rest of the
elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
us the hypergraph isomorphism $\phi$.  The following lemma formalizes
the characeterization of feasible path labeling.

@@

\subsection{ICPPL when given tree is a path}
\label{subsec:icpplicpia}
\xnoindent
Consider a special case of ICPPL with the following properties.
\begin{enumerate}
\item Given tree $T$ is a path. Hence, all path labels are interval labels.
\item Only pairwise intersection cardinality
  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
\item The filter algorithms do not have {\em \bf exit} statements.
\end{enumerate}
This is called an Intersection Cardinality Preservation Interval
Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
used in the next section for finding tree path labeling from a
$k$-subdivided star due to this graph's close relationship with
intervals.

>>>>

\noindent 
\begin{tabbing}
xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
\textbf{COP}   \> Consecutive-ones Property \\
\textbf{COT}   \> Consecutive-ones Testing \\
\textbf{ICPIA}   \> Intersection Cardinality Preservation Interval Assignment \\
\textbf{ICPPL}   \> Intersection Cardinality Preserved Path Labeling \\
% ADD MORE HERE
\end{tabbing}

@@

\tnote{ The main themes of the thesis that must not be left out:\\
  connection of COP to *SET SYSTEMS* and thus to general
  data bases. cite:nsnrs09\\
  *GRAPH ISOMORPHSIM* by canonization cite:kklv10\\
  *CERTIFICATE FOR A PROBLEM* cite:mcc04\\
  Also see purple book notes.}

@@

\section{Consecutive-ones Property Testing - a Brief Survey} %{Background}
\label{sec:background}
%\vspace{\secfirstparatrim}
In this section, a brief survey of the consecutive-ones problem and its
optimization problems is presented.

@@

\section{Application of COP in areas of Graph theory and algorithms}
\subsection{COP in relational database model}
\tnote{(set systems theme)}
\subsection{COP in graph isomorphism}
\tnote{(canonization theme)}
\subsection{Invention of certifying algorithms using COP}
\tnote{ (certification McC04 theme)}

@@

% \subsection{Overlap Graphs and Marginal Hyperedges} 
% An {\em
%   overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph such
% that its vertex set has a bijection to $\cF$ and there exists an edge
% between two of its vertices \iff their corresponding hyperedges
% overlap. Two hyperedges $S$ and $S'$ are said to {\em overlap},
% denoted by $S \overlap S'$, if they have a non-empty intersection and
% neither is contained in the other; $S \overlap S' \text{ iff } S \cap
% S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$
% is a spanning subgraph of $\bI(\cF)$ and not necessarily
% connected. Each connected component of $\bO(\cF)$ is called an {\em
%   overlap component}.
% % If there are $d$ overlap components in $\bO(\cF)$, the set
% % subsystems are denoted by $\cO_1, \cO_2, \ldots \cO_d$. Clearly
% % $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$, it can be
% % verified that one of the following is true.
% % \begin{enumerate}[a) ]
% % \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
% % \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
% % \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
% % \end{enumerate}
% % \vspace{\prelimspace} 
% A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
% \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
% \cite{kklv10}. Additionally, if $S$ is such that it is contained in no
% other marginal hyperedge in $\cF$, then it is called {\em
%   super-marginal}.
% % i.e., it is inclusion maximal then it is called {\em
% %   super-marginal}.

@@

\chapter{Consecutive-ones Property - a Survey of Important Results}
\tnote{TBD: have a few lines about organization of chapter}

@@

\tnote{ADD NEW SECTIONS - MATCHING WHAT'S IN INTRO CHAPTER}

\chapter{Preliminaries to new results}
\label{ch:prelims} 
%\section{Preliminaries} 
%\label{sec:prelims}

\tnote{TBD: have a few lines about organization of chapter}This
section states definitions and basic facts necessary in the scope of
this document.

\tnote{Move some of this to general prelim in chapter 1}

@@

The graph $T$ represents a {\em target tree} with same
number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.
%\vspace{\prelimspace}

% 
% A set system $\cF$ can be alternatively represented by a {\em
%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and hyperedges
% are the sets in $\cF$. This is a known representation for interval
% systems in literature \cite{bls99,kklv10}.  We extend this definition here to
% path systems.

A set system $\cF$ can be alternatively represented by a {\em
  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
are the sets in $\cF$. This is a known representation for interval
systems in literature \cite{bls99,kklv10}.  We extend this definition here to
path systems. Due to the equivalence of set system and hypergraph in
the scope of this paper, we drop the
subscript $_H$ in the notation and refer to both the structures by $\cF$.
%\vspace{\prelimspace}

Two hypergraphs $\cF'$, $\cF''$ are said to be {\em
  isomorphic} to each other, denoted by $\cF' \cong \cF''$, iff there
exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such
that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in
$\cF'$ iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x
\in A\}$ \cite{kklv10}. This is called {\em hypergraph
  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
  required }
%\vspace{\prelimspace}

The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
there exists an edge between two vertices iff their corresponding
hyperedges have a non-empty intersection \cite{mcg04}.
%\vspace{\prelimspace}

@@

If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$
is called a {\em path hypergraph} and $\cP$ is called {\em path
  representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
\rightarrow V(T)$, then it is clear that there is an {\em induced path
labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
$\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
$supp(\cP) = V(T)$.
%\vspace{\prelimspace}

@@

A path labeling $(\cF, \cl)$ is defined to be {\em
  feasible} if
$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
%\vspace{\prelimspace}

An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
a graph such that its vertex set has a bijection to $\cF$ and there
exists an edge between two of its vertices iff their corresponding
hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
  overlap}, denoted by $S \overlap S'$, if they have a non-empty
intersection and neither is contained in the other; $S \overlap S'
\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
necessarily connected. Each connected component of $\bO(\cF)$ is
called an {\em overlap component}.
% If there are $d$ overlap components
% in $\bO(\cF)$, the set subsystems are denoted by $\cO_1, \cO_2, \ldots
% \cO_d$. Clearly $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$,
% it can be verified that one of the following is true.
% \begin{enumerate}[a) ]
% \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
% \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
% \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
% \end{enumerate}
%\vspace{\prelimspace}

A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
{\em super-marginal}.

A {\em star} graph is a complete bipartite graph $K_{1,p}$
which is clearly a tree and $p$ is the number of leaves. The vertex
with maximum degree is called the {\em center} of the star and the
edges are called {\em rays} of the star. A {\em $k$-subdivided star}
is a star with all its rays subdivided exactly $k$ times. The
definition of a {\em ray of a $k$-subdivided star} is extended to the path
from the center to a leaf. It is clear that all rays are of length $k+2$.

@@

\tnote{TBD: have a few lines about organization of chapter}
The necessary preliminaries with definitions
etc. are presented in
Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
characterization of a feasible path labeling and finally,
Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
find the tree path labeling of a given set system from a given
$k$-subdivided tree.

@@

In the remaining part of this section we show that $(\cF,
\cl)$ is feasible if and only if it is an ICPPL and
Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two levels of
filtering of $(\cF, \cl)$ to make it simpler while retaining the set
of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
% One direction of this claim isclear: that if a path
% labeling is feasible, then all intersection cardinalities are
% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL iff input PL
%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
% the sets map to paths, ``preserving'' the path labeling.  The rest
% of the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
its correctness.  This algorithm refines the path labeling by
processing pairs of paths in $\cF^\cl$ that share a leaf until no two
paths in the new path labeling share any leaf.

@@

 \begin{enumerate}[\xbullet]
  \item {\em Invariant} I/II
    \begin{enumerate}[{I/II}a $|$] % \textbullet 
    \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
      trivially true by induction hypothesis.
    \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
      $\cF_{j-1}$, then it must be one of the new sets added in
      $\cF_j$. In this case, it is clear that for each new set, the
      image under $\cl_j$ is a path since by definition the chosen
      sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
      condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
      common leaf. Thus invariant I is proven.\\
      Moreover, due to induction hypothesis of invariant III and the
      definition of $l_j$ in terms of $l_{j-1}$, invariant II is
      indeed true in the $j$th iteration for any of the new sets.  If
      $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
      \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
      If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
      |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
      \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
      |\cl_j(S_1 \setminus S_2)|
      = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
    \end{enumerate}
  \item {\em Invariant} III
    \begin{enumerate}[{III}a $|$]
    \item {\em $R$ and $R'$ are not new sets.} It is in
      $\cF_{j-1}$. Thus trivially true by induction hypothesis.
    \item {\em Only one, say $R$, is a new set.} Due to invariant IV
      induction hypothesis, Lemma~\ref{lem:setminuscard} and
      definition of $\cl_j$, it follows that invariant III is true no
      matter which of the new sets $R$ is equal to. If $R = S_1 \cap
      S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
      \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
      \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
      \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
      |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
      |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
      \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
      S_1$. Note $R'$ is not a new set.\\

@@

As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
is such that there is exactly one set in $\cF$ with $v$ as a vertex in
the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
elements in $supp(\cF)$ whose images are leaves in a hypergraph
isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
  \ne S} \hS$. In other words, $x$ is an element present in no other
set in $\cF$ except $S$. This is intuitive since $v$ is present in no
other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
leaf $v$ are then removed from the set $S$ and path $\cl(S)$
respectively. After doing this for all leaves in $T$, all path images
in the new path labeling $\cl_{new}$ except leaf labels (a path that
has only a leaf is called the {\em leaf label} for the corresponding
single element hyperedge or set) are paths from a new pruned tree $T_0
= T \setminus \{v \mid v \text{ is a leaf in }
T\}$. Algorithm~\ref{leafasgn} is now presented with details.

@@

Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
loop. This will abort our procedure of finding the hypergraph
isomorphism. The following lemma shows that this will not happen.

@@

We have seen two filtering algorithms above, namely,
Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
{\tt filter\_2} which when executed serially repectively result in a
new ICPPL on the same universe $U$ and tree $T$. We also proved that
if the input is indeed feasible, these algorithms do indeed output the
filtered ICPPL. Now we present the algorithmic characterization of a
feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.

Algorithm~\ref{al:icppl-find-isomorph} computes a
hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
and Algorithm~\ref{leafasgn} and pruning the leaves of the input
tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
leaf labels are removed from the path labeling $\cl_2$ and the
corresponding elements are removed from the corresponding sets in
$\cF_2$. This tree pruning algorithm is recursively called on the
altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
recursive call returns the bijection $\phi''$ for the rest of the
elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
us the hypergraph isomorphism $\phi$.  The following lemma formalizes
the characeterization of feasible path labeling.

@@

\subsection{ICPPL when given tree is a path}
\label{subsec:icpplicpia}

Consider a special case of ICPPL with the following properties.
\begin{enumerate}
\item Given tree $T$ is a path. Hence, all path labels are interval labels.
\item Only pairwise intersection cardinality
  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
\item The filter algorithms do not have {\em \bf exit} statements.
\end{enumerate}
This is called an Intersection Cardinality Preservation Interval
Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
used in the next section for finding tree path labeling from a
$k$-subdivided star due to this graph's close relationship with
intervals.
